Parametric analysis of ambisonic audio

### 2020-07-05 Chapters/Conclusions.tex
tbc

The scientific objectives of this thesis, as they were described in Section \ref{sec:objectives}, are:

>>>>

In this thesis we have presented our contributions to different components of an ambisonics analysis and generation framework, with a focus on reproducibility and portability to real-world scenarios.
The main scientific objectives of this thesis, as they were described in Section \ref{sec:objectives}, are:

### 2020-07-05 Chapters/DataGeneration.tex
on work
tbc

\section{MASP: a Python library for multichannel acoustic signal processing}

\subsection{Description}

@@

\textit{pysofaconventions} has gained a moderate amount of attraction. Among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}. Furthermore, given the open nature of the project, several external researchers have contributed with bug fixes or new feature implementations.

@@

\subsection{Experiment reproducibility}

@@

>>>>

\section{MASP: a Python library for multichannel acoustic signal processing}
\label{sec:masp}
\subsection{Description}

@@

%\textit{pysofaconventions} has gained a moderate amount of attraction. Among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}. Furthermore, given the open nature of the project, several external researchers have contributed with bug fixes or new feature implementations.

@@

\subsection{Experiment reproducibility}
\label{experiment_reproducibility}

@@

\section{Conclusion}

The present Chapter has presented the main contributions of the Thesis regarding the availability of ambisonic data, specifically targeting software packages and libraries developed for the generation and storage of ambisonic sound scenes.

The concept of \textit{research reproducibility} is directly related with these topics. According to \cite{cannam2012sound}, providing just an article as a research outcome would not be sufficient for reproducing the research in a feasible way. Instead, supplementary material, such as the code implementation, the dataset, and instructions on how to set up the environment, should be also conveniently provided.

In this regard, the software discussed in this Chapter try to foster reproducible methods and procedures in several different ways.
First, all code is freely available online with open source licenses. 
Second, the contributed software is completely written in Python, which is itself an open-source project. Actually, all code used in this follows the same path, with with the exception of the particle tracker used in Chapter~\ref{chap:seld2019}. 
And third, the motivation of the Ambiscaper project itself is the reproducibility of experimental conditions for signal processing methods using ambisonic data, as explained in detail in Section~\ref{experiment_reproducibility}.\\

Regarding the lifecycle of the libraries described in this Chapter, their status is diverse. Although \textit{pysofaconventions} has been recently published in an academic context (AES Convention in May 2020), it has been available online for more than a year. In fact, during that time it has gained a moderate amount of attraction: among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}. Furthermore, given the open nature of the project, several external researchers have contributed with bug fixes or new feature implementations.

The \textit{masp} library has been also presented in the same context. However, in this case, the library has not been available beforehand, so the interest generated in the community has been moderate.

For its part, \textit{Ambiscaper} has not been able to attract the community in a significant way. This might be due to a number of reasons, including not enough dissemination among potentially interested communities, or a lack of a paradigmatic use-case situation. 
On the contrary, the library that inspired this work, \textit{Scaper}, has been used for the creation of the dataset in Task 4 of DCASE 2019 and 2020 (sound event detection in domestic environments). This situation generates a feedback loop for the maintenance and improvement of the software, motivated by requests from the community.
Given this situation, the future of \textit{Ambiscaper} is linked to the evolution of the research topics and tools in the communities which miggt be potentially interested in using it.

### 2020-07-05 Chapters/Introduction.tex
tbc

Nevertheless, the greatest contributor to the current interest in ambisonics has been the rise of Virtual Reality (VR) in recent years.
Although VR focuses primarily on visual cues, the immersive experience might be greatly enhanced by spatial audio \cite{begault20003}	. In this context, Ambisonics has been rapidly adopted as \textit{de facto} standard for spatial audio transmission, supposedly due to a variety of factors:

@@

  \item [Recording device independence] Regardless of the specific characteristics of an ambisonic microphone, the recorded signal is usually converted into B-Format, which effectively represents the standard exchange format.

@@

Coming back to the Virtual Reality popularity, the following three events might be representative of the growth that it experienced in the second half of the 2010s: 
(i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
(ii) the Time cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
and (ii) M. Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.

\begin{figure}[t!]
  \includegraphics[width=\textwidth]{Figures/Introduction/num_mics_ticks.png}
  \caption{Number of ambisonic microphones released in last years (from \cite{List_of_Ambisonic_hardware}). From left to right, the vertical lines correspond to (i) Oculus acquisition, (ii) Time cover page on VR, (iii) M. Zuckerberg's speech in MWC, and (iv) Jaunt announcement of shift towards AR.}
  \label{fig:nummics}
\end{figure}

@@

At the present time, however, the high expectations in VR have significantly lowered, as shown in Figure~\ref{fig:vrfunding}. This is due to a variety of reasons, including lack of interesting content and the high production cost of the headsets \cite{fortune}. The change of focus of Jaunt (formerly one of the biggest VR film production companies) towards Augmented Reality (AR), as of October 2018, might be a paradigmatic example of this tendency \cite{theverge}.

@@

B-Format analysis has been divided into three different categories:
\begin{description}
	\item[Acoustic Parameter Estimation] Low-level analysis of the sound field, which yields quantitative information about different acoustic parameters used to model the acoustic scene.
	\item[Signal Enhancement] Modification of the sound field in order to obtain a set of output signals with desired attributes. 
	\item[Scene Description] Textual representation of different high-level characteristics of the sound field.
\end{description}

@@

There are a variety of well known signal enhancement problems, including dereverberation \cite{braun2018speech}, source separation \cite{gannot2017consolidated}, or foreground-background segmentation \cite{stefanakis_foreground_2015}. 
As it has been shown, many of them benefit (or even depend) from the knowledge derived by the acoustic parameter estimation step.

Finally, under the scene description typology we can find a set of applications that provide abstract representations of the sound scene under analysis. Most of the recent research performed in this scope is grouped around the Detection and Classification of Acoustic Scenes and Events (DCASE) community\footnote{\url{http://dcase.community/community_info}}. 
Examples of the problems under consideration are acoustic scene classification \cite{Mesaros2018_DCASE}, sound event localization and detection \cite{Adavanne2018_JSTSP} or audio captioning \cite{drossos2017automated}.

@@

\begin{enumerate}
	\item To develop methods for the characterisation of acoustic parameters from recordings originated from ambisonic microphones. 
	\item To propose methodologies for sound event localization and detection in ambisonic domain which are grounded on spatial parametric analysis. 
	\item To contribute in the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
\end{enumerate}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Contributions}

In the following list we show the main scientific contributions of this dissertation, organised by Chapters:

\begin{itemize}

	\item Chapter 3:\\
	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}

	\item Chapter 4:\\
	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}

	\item Chapter 5:\\

	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.

	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.

	\item Chapter 6:\\
	\textbf{"Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes".}\\
	\underline{A. P\'erez-L\'opez}.
	In \textit{16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}

	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
	\underline{A. P\'erez-L\'opez} and J. De Muynke.
	In \textit{Audio Engineering Society Convention 144. Audio Engineering Society, 2018.}

	\textbf{"pysofaconventions, a Python API for SOFA".}
	\underline{A. P\'erez-L\'opez}.
	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}

	\textbf{"A Python library for Multichannel Acoustic Signal Processing".}
	\underline{A. P\'erez-L\'opez} and A. Politis.
	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}

\end{itemize}

Moreover, although not strictly aligned with the research direction of this thesis, the author has collaborated in the following publications:

\begin{itemize}
	\item \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.

\end{itemize}

Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released. All of them are available through the author's GitHub page \cite{andresgithub}:
\begin{itemize}

	\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation}

	\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020}

	\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3}

	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}

	\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions}

	\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}

	\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper}

\end{itemize}

>>>>

Nevertheless, the greatest contributor to the current interest in ambisonics has been the rise of Virtual Reality (VR) in recent years.
Although VR focuses primarily on visual cues, the immersive experience can be greatly enhanced by spatial audio \cite{begault20003}. In this context, Ambisonics has been rapidly adopted as \textit{de facto} standard for spatial audio transmission, supposedly due to a variety of factors:

@@

  \item [Recording device independence] Regardless of the specific characteristics of an ambisonic microphone, the recorded signal is usually converted into B-Format, which is effectively  the standard exchange format.

@@

Coming back to the issue of the popularity of Virtual Reality, we have selected three events that might epitomize the growth undergone in the second half of the 2010s:

%Coming back to the Virtual Reality popularity, the following three events might be representative of the growth experienced in the second half of the 2010s: 
\begin{enumerate}
	\item The billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}
	\item \textit{Time} maagazine cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
	\item  M. Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.
\end{enumerate}

%(i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
%(ii) the Time cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
%and (iii) M. Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.

\begin{figure}[t!]
  \includegraphics[width=\textwidth]{Figures/Introduction/num_mics_ticks.png}
  \caption{Number of ambisonic microphones released in last years (from \cite{List_of_Ambisonic_hardware}). From left to right, the vertical lines correspond to (1) Oculus acquisition, (2) \textit{Time} cover page on VR, (3) M. Zuckerberg's speech in MWC, and (4) Jaunt announcement of shift towards AR.}
  \label{fig:nummics}
\end{figure}

@@

At the present time, however, the high expectations put into VR have significantly lowered, as shown in Figure~\ref{fig:vrfunding}. This is due to a variety of reasons, including lack of interesting content and the high production cost of the headsets \cite{fortune}. The change of focus of Jaunt (formerly one of the biggest VR film production companies) towards Augmented Reality (AR), as of October 2018, might be a paradigmatic example of this tendency \cite{theverge}.

@@

Ambisonic signal analysis has been divided into three different categories:
\begin{description}
	\item[Signal Enhancement] Modification of the input signal in order to obtain one or more output signals with desired attributes.

	There are a variety of well known signal enhancement problems, including dereverberation \cite{braun2018speech}, source separation \cite{gannot2017consolidated}, or foreground-background segmentation \cite{stefanakis_foreground_2015}. 
	As it has been shown, many of them benefit (or even depend) from the knowledge derived by acoustic parameter estimation methods.

	\item[Acoustic Parameter Estimation] Low-level analysis of the sound field, which yields quantitative information about different acoustic parameters used to model the acoustic scene.

@@

	\item[Scene Description] Textual representation of different high-level characteristics of the sound field.

	Under the scene description typology we can find a set of applications that provide abstract representations of the sound scene under analysis. Most of the recent research performed in this scope is grouped around the Detection and Classification of Acoustic Scenes and Events (DCASE) community \cite{dcase}. 
	Examples of the problems under consideration are acoustic scene classification \cite{Mesaros2018_DCASE}, sound event localization and detection \cite{Adavanne2018_JSTSP} or audio captioning \cite{drossos2017automated}.
\end{description}

@@

\begin{enumerate}
	\item To develop methods for the characterisation of acoustic parameters from recordings originated from ambisonic microphones. 
	\item To propose methodologies for sound event localization and detection in ambisonic domain which are grounded on spatial parametric analysis. 
	\item To contribute to the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
\end{enumerate}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Contributions}
%
%
%In the following list we show the main scientific contributions of this dissertation, organised by Chapters:
%
%\begin{itemize}
%
%	\item Chapter 3:\\
%	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
%	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
%	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}
%	
%	\item Chapter 4:\\
%	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
%	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
%	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
%
%	\item Chapter 5:\\
%	
%	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
%	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
%	
%	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
%	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
%	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
%	
%	
%	\item Chapter 6:\\
%	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
%	\underline{A. P\'erez-L\'opez}.
%	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
%	
%	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
%	\underline{A. P\'erez-L\'opez} and J. De Muynke.
%	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}
%	
%	\textbf{"pysofaconventions, a Python API for SOFA".}
%	\underline{A. P\'erez-L\'opez}.
%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}
%	
%	\textbf{"A Python library for multichannel acoustic signal processing".}
%	\underline{A. P\'erez-L\'opez} and A. Politis.
%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}
%	
%\end{itemize}
%
%Moreover, although not strictly aligned with the research direction of this thesis, the author has collaborated in the following publications:
%
%\begin{itemize}
%	\item \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
%	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
%	
%\end{itemize}
%
%
%Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released. All of them are available through the author's GitHub page \cite{andresgithub}:
%\begin{itemize}
%
%	\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation}
%	
%	\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020}
%	
%	\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3}
%	
%	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
%
%	\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions}
%	
%	\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	
%
%	\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} 
%	
%
%
%\end{itemize}

### 2020-07-05 Chapters/ReververationTimeEstimation.tex
tbc

We have presented in this work a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings. Our method is based on a first dereverberation step,  performed by a multichannel autoregressive model of the late reverberation. The resulting dry signal is then used to estimate the impulse response decay by means of system identification. 
The performance of the method is evaluated in a simulated experimental environment with two different reverberant datasets, and compared against a state-of-the-art method. 
Results show that our method outperforms the baseline method in a majority of evaluation metrics and conditions, and consistently provides results with less variability than the baseline method. 
In future work, we plan to extend the experimental setup by using recorded IRs. Furthermore, the proposed method could be extended to the case of moving sources by using an \textit{online} autoregressive model. 
%Finally, an evaluation of the method in terms of the number of channels or the ambisonic order remains to be done.
Finally, an extension of the method for higher ambisonic orders remains to be done.

>>>>

I n this Chapter, we have presented  a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings. Our method is based on a first dereverberation step,  performed by a multichannel autoregressive model of the late reverberation. The resulting dry signal is then used to estimate the impulse response decay by means of system identification. 
The performance of the method is evaluated in a simulated experimental environment with two different reverberant datasets, and compared against a state-of-the-art method. 
Results show that our method outperforms the baseline method in a majority of evaluation metrics and conditions, and consistently provides results with less variability than the baseline method. 
In future work, we plan to extend the experimental setup by using recorded IRs. Furthermore, the proposed method could be extended to the case of moving sources by using an \textit{online} autoregressive model. 
%Finally, an evaluation of the method in terms of the number of channels or the ambisonic order remains to be done.
Finally, an extension of the method for higher ambisonic orders remains to be done.
